Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-02-28 paper/main.tex
lemma in section 3.1 proven!

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
  only minimal requirements on $\Te$. It does not use the
  information about automaton structure, but it comes with an
  interesting simplification of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is a balanced version. It says that every
  transition in $\Te$ has to have a preceding and a succeeding
  transition in $\Te$, which is a necessary condition on cycles, but
  not a sufficient one.
\item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
  cycle. Unfortunately, it uses additional universally quantified
  variables corresponding to automata states. Transformation to prenex
  normal form turns the quantifiers to existential ones and the
  resulting formula thus contains quantifier alternation.
\end{itemize}

\subsection{Lightweight version $\mcycle_1(\vec{e})$}
The lightweight version is based on the fact that every cycle is nonempty.
Indeed, the only restriction imposed by
\[
  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
\]
is that $\Te$ is nonempty.

@@

Formally, we define an equivalence on transitions such that for each
$t_1,t_2\in\delta$ we write $t_1\sim t_2$ whenever
$\mks(t_1)=\mks(t_2)$.
%This relation is clearly an equivalence.
We prove the following statement. 
\begin{lemma}
  Assume that there is a function $\rem$ and a formula $\psi$ such
  that for every nonempty set $T\subseteq\delta$ it holds that $T$
  satisfies $\varphi$ iff $\rem(T)$ satisfies $\psi$. Then we can
  safely assume that $\rem$ respects the quivalence $\sim$, that is
  $t_1\sim t_2\implies\rem(t_1)=\rem(t_2)$.
\end{lemma}
\begin{proof}
  \todo{Jan is here!}
\end{proof}

Formally, for each pair of transitions $t_1,t_2\in\delta$, we write
$t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$. This relation is clearly an equivalence.
Further, let us say that

We show that if this formula holds for some $\rem$ and $\psi$,
then it also holds for some $\rem'$ and $\psi$ such that $\rem'$
assigns the same sets of acceptance marks to transitions originally
marked by the same sets of marks.

>>>>

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
  only minimal requirements on $\Te$. It does not use the
  information about automaton structure, but it comes with an
  interesting simplification of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is a balanced version. It says that $\Te$
  is nonempty and every transition in $\Te$ has to have a preceding
  and a succeeding transition in $\Te$, which is a necessary condition
  for being a cycle, but not a sufficient one.
\item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
  cycle. Unfortunately, it uses additional universally quantified
  variables corresponding to automata states. Transformation to prenex
  normal form turns the quantifiers to existential ones and the
  resulting formula thus contains quantifier alternation.
\end{itemize}

\subsection{Lightweight version $\mcycle_1(\vec{e})$}
The lightweight version is defined as
\[
  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
\]
which means only that $\Te$ is nonempty. This condition is satisfied by
every cycle.

@@

Formally, we define an equivalence $\sim\subseteq\delta\times\delta$
on transitions such that $t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$.
\begin{lemma}
  Assume that there is a function $\rem$ and a formula $\psi$ such that 
  \[
    \textrm{for every nonempty set $T\subseteq\delta$ it holds }
      (T~\msat~\varphi~\iff~\rem(T)~\msat~\psi).\tag{$1$}\label{eq:lem1} 
  \]  
  Then there exists a function $\rem'$ that respects % the equivalence
  $\sim$ (i.e.,
  if $t_1\sim t_2$ then $\rem'(t_1)=\rem'(t_2)$) and
  \[
    \textrm{for every nonempty set $T\subseteq\delta$ it holds }
    (T~\msat~\varphi~\iff~\rem'(T)~\msat~\psi).\tag{$2$}\label{eq:lem2} 
  \]  
\end{lemma}
\begin{proof}
  Let $\rem$ be a function and $\psi$ a formula such that
  \eqref{eq:lem1} holds. To construct the function $\rem'$, we first
  select one transition from each equivalence class of $\sim$. For
  every transition $t\in\delta$, by $\overline{t}$ we denote the
  selected transition equivalent to $t$. We define the function
  $\rem'$ as $\rem'(t)=\rem(\overline{t})$. Clearly, $\rem'$ respects
  the equivalence $\sim$. It remains to show that \eqref{eq:lem2}
  holds for $\rem'$ and $\psi$.

  Let $T\subseteq\delta$ be an arbitrary nonempty set. We construct
  the set $\overline{T}=\{\overline{t}\mid t\in T\}$. As
  $\mks(t)=\mks(\overline{t})$ holds for all transitions, we get
  $\mks(T)=\mks(\overline{T})$ and thus
  \[
    T~\msat~\varphi~\iff~\overline{T}~\msat~\varphi.
  \]
  Now we apply $\eqref{eq:lem1}$ on $\overline{T}$ to get that
  \[
    \overline{T}~\msat~\varphi~\iff~\rem(\overline{T})~\msat~\psi.
  \]
  Finally, the definition of $\rem'$ implies that
  $\rem'(T)=\rem(\overline{T})$ and thus
  \[
    \rem(\overline{T})~\msat~\psi~\iff~\rem'(T)~\msat~\psi.
  \]
  Altogether, we obtain
  \[
    T~\msat~\varphi~\iff~\overline{T}~\msat~\varphi~\iff~\rem(\overline{T})~\msat~\psi~\iff~\rem'(T)~\msat~\psi
  \]
  which proves that \eqref{eq:lem2} holds for $\rem'$ and $\psi$. 
\end{proof}

The previous lemma and its proof suggest implies that it is sufficient to
